---
title: "Eurostat EDA"
author: "Sindre H. Øveraas, Alen Colakovic & Sebastian M. Fløysand"
format: pdf
editor: visual
bibliography: references.bib
---

## Eurostat EDA

### Sub-National GDP

<<<<<<< HEAD
# Meta Data Description

### Nama 10r

Regional accounts are based on the same definitions and concepts as national accounts, this is because they are a specification of the national accounts.

*Gross domestic product (GDP) is the standard measure of the value added created through the production of goods and services in a country during a certain period. As such, it also measures the income earned from that production, or the total amount spent on final goods and services (less imports). While GDP is the single most important indicator to capture economic activity, it falls short of providing a suitable measure of people's material well-being for which alternative indicators may be more appropriate.\
*\@GDPSpendingGross

We have two ways of calculating this within regional accounts:

**The first one is the output approach**

*Where the GDP is the sum of gross value added of the various institutional sectors or the various industries plus taxes and less subsidies on products. It is also the balancing item in the total economy production account."*

**2. Income approach**

*"GDP is the sum of uses in the total economy generation of income account: compensation of employees plus gross operating surplus and mixed income plus taxes on products less subsidies plus consumption of fixed capital.*

*Contrary to national accounts GDP is not compoled from the expenditure side in regional accounts due to data limitations on the inter-regional flows of goods and services.*

*The different measures for the regional GDP are absolute figures in € and Purchasing Power Standards (PPS), figures per inhabitant and relative data compared to the EU Member States average."*

@RegionalEconomicAccounts
=======
Country codes: BE - Belgium, BG - Bulgaria, HR - Croatia, IT - Italy, AT - Austria, SE - Sweden, RS - Serbia

```{r, message=FALSE}
library(tidyverse)
library(vtable)
library(dineq)
library(dplyr)
```

```{r}
library(readr)
GDP <- read.csv('nama_10r_3gdp__custom_3564935_linear.csv')
Population <- read_csv("demo_r_pjanaggr3__custom_3579517_linear.csv")

gdpdata <- GDP %>%
  rename(Year = TIME_PERIOD, GDP = OBS_VALUE, Region = geo)

populationdata <- Population %>%
  rename(Year = TIME_PERIOD, Population = OBS_VALUE, Region = geo)

GDP_Per_Capita <- gdpdata %>%
  left_join(populationdata, by=c("Region", "Year")) %>%
  select(Region, Year, GDP, Population) %>%
  mutate(
    GDP_capita = (GDP * 1000000)/Population
  )
```

```{r, echo=FALSE}
summary(GDP_Per_Capita[,3:5])
```

```{r}
GDP_Per_Capita$GDP_capita <- as.numeric(GDP_Per_Capita$GDP_capita)
GDP_Per_Capita$Population <- as.numeric(GDP_Per_Capita$Population)
```

```{r}
gini.wtd(GDP_Per_Capita$GDP_capita, weights = GDP_Per_Capita$Population)
```

```{r}
GDP_Per_Capita <- GDP_Per_Capita %>%  
  mutate(NUTS2 = substr(GDP_Per_Capita$Region,1,4))

GDP_Per_Capita <- GDP_Per_Capita %>%  
  mutate(NUTS = substr(GDP_Per_Capita$Region,1,2))

GDP_Per_Capita  %<>%
  group_by(NUTS2) %>%
  na.exclude(GDP_Per_Capita) %>%
  mutate(GINI_NUTS2 = gini.wtd(GDP_capita, weights = Population)) %>%
  ungroup()
```

```{r}
summary(GDP_Per_Capita[8])
```

```{r, echo=FALSE}
ggplot(data = GDP_Per_Capita) + geom_density(aes(x = GINI_NUTS2), fill = "maroon", col = "maroon", alpha = 0.2) + theme_bw()
```

```{r, echo=FALSE}
ggplot(data = GDP_Per_Capita) + geom_histogram(aes(x = GINI_NUTS2), fill = "maroon", col = "maroon", alpha = 0.2) + theme_bw()
```

Looking at the plot above there is one outlier up against 0.4 with around 100 observations.
The same result also seem to accure in the density plot.
>>>>>>> Sindre
